CGAT In The Hat

### 2012-12-06 paper/paper.tex
Added first version of Database choice section.

\section{Motivation}\label{sec:motive}
Dr. Anya Goodman, a professor in the biochemistry department at Cal Poly, San Luis
Obispo, offers a bioinformatics course that covers several aspects of gene
annotation. Her work is a part of the Genomics Education Partnership (GEP) at 
Washington University in St. Louis.\cite{gep} The goal of GEP is to teach undergrad students
in Biology how to annotate genes and other features in raw sequence data.
The CGAT project is spearheaded by Ryan Verdon to accommodate Dr.
Goodman's needs and ideas for ideal gene annotation software.

\section{Database Choice}\label{sec:choice}
% TODO(ryan): Put decision and considereations here.
%  Schema will go in implementation section.
%  Incoude some text about how having the documents in JSON (Mongo is actually BSON) makes it really easy to use in JSON.

>>>>

\section{Motivation}\label{sec:motive}
Dr. Anya Goodman, a professor in the Chemistry department at Cal Poly, San Luis
Obispo, offers a bioinformatics course that covers several aspects of gene
annotation. Her work is a part of the Genomics Education Partnership (GEP) at 
Washington University in St. Louis.\cite{gep} The goal of GEP is to teach undergrad students
in Biology how to annotate genes and other features in raw sequence data.
The CGAT project is spearheaded by Ryan Verdon to accommodate Dr.
Goodman's needs and ideas for ideal gene annotation software.

\section{Database Choice}\label{sec:choice}
The choice for backend database was between MySQL, Couchbase, and MongoDB.
We included MySQL in the options because its the standard that small websites tend to use. Then we included both
Couchbase and MongoDB in our considerations because they are popular NoSQL database that are different types of NoSQL
databases. Couchbase is a key-value store unlike MongoDB which is a document-store.

We ended up using MongoDB for several reasons. Firstly, the document structure was extremely helpful when storing
relation. We ended up attaching a lot of data on our relations that would of been a pain to store in SQL but being
able to store the data as complex objects in JSON was super helpful. Secondly, being able to store the data in JSON
(BSON in MongoDB) made it extremely easy to retrieve and use the data in a web application. Because of these two reasons
we narrowed our choices down to MongoDB and Couchbase.

When we were deciding between MongoDB and Couchbase we looked at the features they 
provided and noticed several things. First, Couchbase currently is a pure key-value store it has
no idea what the structure of the value. Unlike MongoDB, which can query for fields inside an 
object specified by a key. This feature of MongoDB makes updating relations (which we 
have a lot of) super easy. Second, Couchbase provides a huge
cluster/map reduce overhead that is not necessary for our data.
